---
title: "Clean income data"
output:
  html_document:
    df_print: paged
---

This notebook produces a few easier-to-use datasets from the input data from the IDI.
This input data *contains no total rows*, only strict aggregates of income bracket by age bracket by post-school field of study by qualification level by employment status. 

```{r}
library(tidyverse)
library(readxl)
library(openxlsx)
library(DBI)
source("../scripts/unz_common.R")
```

```{r}
data <- read_excel("../input_data/income_wide_2018.xlsx", na = "NA")
head(data)
```

Combining level 5 and 6 diplomas

```{r}
level5_dips <- filter(data, `Highest Qualification` == "Level 5 Diploma")
level6_dips <- filter(data,`Highest Qualification` == "Level 6 Diploma")
dips <- merge(level5_dips, level6_dips, by = c("Work Status", "Post School Qualification Field of Study", "Age Group"), all = TRUE)

dips[is.na(dips)] <- 0
dips$`Highest Qualification.x` <- "Level 5 Diploma"
dips$`Highest Qualification.y` <- "Level 6 Diploma"
dips <- mutate(dips, "Highest Qualification"="Total 5 & 6 Diploma")
dips <- mutate(dips, Loss=Loss.x+Loss.y)
dips <- mutate(dips, `Zero income`=`Zero income.x`+`Zero income.y`)
dips <- mutate(dips, `$1-$5,000`=`$1-$5,000.x`+`$1-$5,000.y`)
dips <- mutate(dips, `$5,001-$10,000`=`$5,001-$10,000.x`+`$5,001-$10,000.y`)
dips <- mutate(dips, `$10,001-$15,000`=`$10,001-$15,000.x`+`$10,001-$15,000.y`)
dips <- mutate(dips, `$15,001-$20,000`=`$15,001-$20,000.x`+`$15,001-$20,000.y`)
dips <- mutate(dips, `$20,001-$25,000`=`$20,001-$25,000.x`+`$20,001-$25,000.y`)
dips <- mutate(dips, `$25,001-$30,000`=`$25,001-$30,000.x`+`$25,001-$30,000.y`)
dips <- mutate(dips, `$30,001-$35,000`=`$30,001-$35,000.x`+`$30,001-$35,000.y`)
dips <- mutate(dips, `$35,001-$40,000`=`$35,001-$40,000.x`+`$35,001-$40,000.y`)
dips <- mutate(dips, `$40,001-$50,000`=`$40,001-$50,000.x`+`$40,001-$50,000.y`)
dips <- mutate(dips, `$50,001-$60,000`=`$50,001-$60,000.x`+`$50,001-$60,000.y`)
dips <- mutate(dips, `$60,001-$70,000`=`$60,001-$70,000.x`+`$60,001-$70,000.y`)
dips <- mutate(dips, `$70,001-$100,000`=`$70,001-$100,000.x`+`$70,001-$100,000.y`)
dips <- mutate(dips, `$100,001-$150,000`=`$100,001-$150,000.x`+`$100,001-$150,000.y`)
dips <- mutate(dips, `$150,001 or More`=`$150,001 or More.x`+`$150,001 or More.y`)

names(dips)
combined_dips_final <- dips[,c(1,38,2,3,39:54)]

rbind(data, combined_dips_final)
```

Adding code variables for field and study level
```{r}
data2 <- data %>% 
    append_narrow_field_of_study_code(field_var = "Post School Qualification Field of Study", code_var = "field_code") %>%
    append_study_level_code(level_var = "Highest Qualification", code_var = "level_code")
head(data2)
```



Define standard data used (rows corresponding to the standard qualifications that we are interested in). The helps prevent double counts.
```{r}
data3 <- data2 %>%
    mutate(
#        set_total = `Highest Qualification` == "Total" | `Post # School Qualification Field of Study` == "Total",
        set_standard = `Highest Qualification` %in% 
            c('No Qualification', 'Level 1 Certificate', 'Level 2 Certificate', 'Level 3 Certificate', 
              'Level 4 Certificate', 'Total 5 & 6 Diploma', 'Bachelor Degree and Level 7 Qualification', 
              'Post-graduate and Honours Degrees', 'Masters Degree', 'Doctorate Degree', 
              'Overseas Secondary School Qualification') 
        
    ) %>% mutate(
        #Fulltime code = 1, parttime = 3, all = 2
        employed_code = ifelse(`Work Status` == 'Employed full-time', 1, ifelse(`Work Status` == 'Total employed', 2, 3))
    )

head(data3)
```

Rename columns so that there are not spaces in the column names.
```{r}
cen_income_bands <- data3 %>%
    rename(
        level = "Highest Qualification", 
        field = "Post School Qualification Field of Study", 
        people_count = "Total", 
        mean_income = "Mean Income ($)",
        median_income = "Median Income ($)",
        age = "Age Group",
        work_status = `Work Status`
    ) %>% select(
        set_total, set_standard, employed_code, work_status, level_code, level, field_code, field, age, everything() 
    )

cen_income_bands
```


A connection to an analysis database
```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

Write the table before any transformations are done.
```{r}
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "cen_income_bands"), cen_income_bands, overwrite = TRUE)
```

We take the sum of the product of each count and income band midpoint and then divide by the number of people to ge the mean income. 
We do this for two distinct populations:
'Restricted' counts and incomes exclude those who earned $25,000 a year or less. Counts and incomes with suffix '2' exclude those who earned no income or made a loss.

```{r}
cen_annual_avg_income <- cen_income_bands %>% mutate(
    
    restricted_people_count = `$25,001-$30,000` + `$30,001-$35,000` + `$35,001-$40,000` + 
                              `$40,001-$50,000` + `$50,001-$60,000` + `$60,001-$70,000` +
                              `$70,001-$100,000` + `$100,001-$150,000` + `$150,001 or More`,
    
    restricted_total_income = `$25,001-$30,000` * 27500 + `$30,001-$35,000` * 32500 + `$35,001-$40,000` * 37500 + 
                              `$40,001-$50,000` * 45000 + `$50,001-$60,000` * 55000 + `$60,001-$70,000` * 65000 +
                              `$70,001-$100,000` * 85000 + `$100,001-$150,000` * 125000 + `$150,001 or More` * 150000,
    
    people_count2           = `$1-$5,000` + `$5,001-$10,000` +
                              `$10,001-$15,000` + `$15,001-$20,000` + `$20,001-$25,000` +
                              `$25,001-$30,000` + `$30,001-$35,000` + `$35,001-$40,000` + 
                              `$40,001-$50,000` + `$50,001-$60,000` + `$60,001-$70,000` +
                              `$70,001-$100,000` + `$100,001-$150,000` + `$150,001 or More`,
    
    total_above_zero_income = `$1-$5,000` * 2500 + `$5,001-$10,000` * 7500 +
                              `$10,001-$15,000` * 12500 + `$15,001-$20,000` * 17500 + `$20,001-$25,000` * 22500 +
                              `$25,001-$30,000` * 27500 + `$30,001-$35,000` * 32500 + `$35,001-$40,000` * 37500 + 
                              `$40,001-$50,000` * 45000 + `$50,001-$60,000` * 55000 + `$60,001-$70,000` * 65000 +
                              `$70,001-$100,000` * 85000 + `$100,001-$150,000` * 125000 + `$150,001 or More` * 200000
    
) %>% mutate(
        restricted_mean_income = restricted_total_income / restricted_people_count
        , mean_income2 = total_above_zero_income / people_count2
    ) %>% 
    select(set_total, set_standard, employed_code, work_status, level_code, level, field_code, field, age, people_count, restricted_people_count, people_count2, mean_income2, mean_income, median_income, restricted_mean_income) %>%
    mutate(
        #Rounding to the nearest hundred
        restricted_mean_income = round2(restricted_mean_income, -2),
        mean_income2 = round2(mean_income2, -2),
        mean_income = ifelse(mean_income == 0, NA, mean_income),
        median_income = ifelse(median_income == 0, NA, median_income),
        restricted_mean_income = ifelse(restricted_mean_income == 0, NA, restricted_mean_income),
        mean_income2 = ifelse(mean_income2 == 0, NA, mean_income2)
    
    )
head(cen_annual_avg_income)
```
Writing table to database
```{r}
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_avg_income"), cen_annual_avg_income, overwrite = TRUE)
```

Adding data for counts and mean incomes limited to those who earned more than $25,000 AND employed full-time.
```{r}
cen_annual_avg_income2 <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_avg_income"))
unique(cen_annual_avg_income2$level_code)
temp <- cen_annual_avg_income2 %>% filter(employed_code == 1) %>%
    select(-c(people_count, mean_income)) %>%
    rename(people_count = restricted_people_count, mean_income = restricted_mean_income) %>%
    mutate(median_income = NA, employed_code = 4, work_status = "Employed full-time (income restricted)")
head(temp)
```

Writing table to database
```{r}
cen_annual_avg_income2_a <- rbind(cen_annual_avg_income2 %>% select(-c(restricted_people_count, restricted_mean_income)), temp)
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_avg_income2"), cen_annual_avg_income2_a, overwrite = TRUE)
```

