---
title: "Income imputation"
output:
  html_document:
    df_print: paged
---
This notebook uses the cen_annual_avg_income dataset and produces a yearly income dataset for the working life of a graduate.

```{r}
library(tidyverse)
library(readxl)
library(openxlsx)
library(DBI)
source("../scripts/unz_common.R")
```

## A connection to an analysis database
```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

Removing total rows and those who did not earn anything or made a loss
```{r}
data <- dbGetQuery(go_analysis_con, "SELECT *
  FROM [lng].[cen_annual_avg_income]
  where set_standard = 1") %>% select(-c(set_standard, work_status, restricted_people_count, restricted_mean_income)) %>%
    rename(people_count = people_count2, mean_income = mean_income2)
head(data)
```

# Calculate total people count for each field/level
Compare to counts in occupation by qual input data file?
```{r}
people_counts <- filter(data, age != 'Total') %>% group_by(employed_code, level_code, field_code) %>% 
  summarise(people_count = sum(people_count)) %>% 
  filter(people_count > 0)
head(people_counts)
```

# Decide if we should impute for each given field/level group?
## count the missing -> not really used?
```{r}
missing_count <- data %>% group_by(employed_code, field, level) %>%
    summarise(missing_count = sum(is.na(mean_income)))
head(missing_count)
```

## convert income data to wide with age as columns:
The age columns must be in sequence for this to work, the index of the age columns are manually defined and will result in highly undesirable behaviour if they move.
```{r}
age_cols <- 6:16

    wide_form <- data %>% select(-c("people_count")) %>% spread(age, mean_income)
    #print(wide_form)
    print(colnames(wide_form))
    colnames(wide_form)[age_cols] <- c('age_15_19', 'age_20_24', 'age_25_29', 'age_30_34', 'age_35_39', 'age_40_44', 'age_45_49', 'age_50_54', 'age_55_59', 'age_60_64', 'age_65_')
    print(colnames(wide_form))
    head(wide_form)
```

## make the call on what can be imputed:
```{r}
should_impute <- wide_form %>% left_join(missing_count, by=c("employed_code","field", "level")) %>% 
    mutate(
           can_impute = !
            (# Can't impute if there is a run of four missing between 20 and 54
                (is.na(age_20_24) & is.na(age_25_29) & is.na(age_30_34) & is.na(age_35_39) & is.na(age_40_44)) |
                (is.na(age_25_29) & is.na(age_30_34) & is.na(age_35_39) & is.na(age_40_44)) |
                (is.na(age_30_34) & is.na(age_35_39) & is.na(age_40_44) & is.na(age_45_49)) |
                (is.na(age_35_39) & is.na(age_40_44) & is.na(age_45_49) & is.na(age_50_54)) |
                # Can't impute if there is a run of three missing between 50 and 65+
                (is.na(age_50_54) & is.na(age_55_59) & is.na(age_60_64)) |
                (is.na(age_55_59) & is.na(age_60_64) & is.na(age_65_))
            )
    ) %>% left_join(people_counts, by=c("employed_code", "field_code", "level_code"))
head(should_impute)
```

# Do imputation:
## prep fuller for age 15-19 by level

#NB: Re, getting imputation incomes for 15-19 year olds by employment status with quals 8 and above by using max incomes for levels < 8: used for calculating lifetime earnings of what people WOULD have been earning.
```{r}
###Getting weighted means by qual level and occupation across fields of study, for levels < 8, using maximum as imputation value for levels >=8
level_15_19_mean_incomes <- data %>% 
    filter(age == "15-19 years" & !is.na(mean_income) & !level_code %in% c('08', '09', '10')) %>% 
    group_by(employed_code, level_code) %>% 
    summarise(weighted_mean_income = weighted.mean(mean_income, people_count), people_count = sum(people_count))
print(level_15_19_mean_incomes)

### imputation value for level 8+
higher_level_early_income_impute <- (level_15_19_mean_incomes %>% summarise(weighted_mean_income = max(weighted_mean_income)))
higher_level_early_income_impute

### assign level 8 - 10 imputation values
higher_level_imputes <- tibble(employed_code = c(rep(1,3), rep(2,3), rep(3,3)), level_code = rep(c('08', '09', '10'),3))
higher_level_imputes <- higher_level_imputes %>% left_join(higher_level_early_income_impute, by="employed_code")
higher_level_imputes

### bind together the level 1 - 6 dataset with level 7 - 10 (rounding earlier < 8 incomes and removing counts)
age_15_19_imputable_income <- rbind(level_15_19_mean_incomes %>% ungroup() %>% select(c("employed_code","level_code","weighted_mean_income")), higher_level_imputes)  %>%
    mutate(mean_income = round2(weighted_mean_income, -2)) %>% select(-c("weighted_mean_income"))
age_15_19_imputable_income

##? join age 15 - 19 income dataset with wide income dataset
ready_to_impute <- should_impute %>% 
    left_join(age_15_19_imputable_income, by=c("employed_code","level_code")) %>%
    rename(age_15_19_impute_income = mean_income)
head(ready_to_impute)
saveRDS(ready_to_impute,paste0("../datasets/dump_pre_impute_ft.rds"))
```

## Do actual imputation:
```{r}
impute_value = function(row) {
    if (row["level_code"] %in% c('07', '08', '09', '10')) {
        row["age_15_19"] <- row["age_15_19_impute_income"] # change these anyway
    }
    
    if (row["can_impute"] == "TRUE" && as.numeric(row["missing_count"]) > 0) {
        # first age group imputation:
        if (is.na(row["age_15_19"])) {
            row["age_15_19"] <- row["age_15_19_impute_income"] 
        } 
        # impute last age group:
        if (is.na(row["age_65_"])) {
            if (!is.na(row["age_60_64"])) {
                row["age_65_"] <- row["age_60_64"]
            } else if (!is.na(row["age_55_59"])) {
                row["age_65_"] <- row["age_55_59"]
            } else {
                warning("To much missing information to impute last value")
            }
            row["age_65_"] <- row["age_15_19_impute_income"] 
        }
        for(age_group in 7:15) {
            if(is.na(row[age_group])) {
                if (!is.na(row[age_group + 1])) {
                    row[age_group] <- mean(as.numeric(row[(age_group - 1)]), as.numeric(row[(age_group + 1)]))
                } else if (age_group + 2 <= max(age_cols) && !is.na(row[age_group + 2])) {
                    row[age_group] <- round2(as.numeric(row[(age_group - 1)]) + (1/3) * as.numeric(row[(age_group + 2)]) - as.numeric(row[(age_group - 1)]),-2) 
                } else if (age_group + 3 <= max(age_cols) && !is.na(row[age_group + 3])) {
                    row[age_group] <- round2(as.numeric(row[(age_group - 1)]) + (1/4) * as.numeric(row[(age_group + 3)]) - as.numeric(row[(age_group - 1)]),-2)
                } else if (age_group + 4 <= max(age_cols) && !is.na(row[age_group + 4])) {
                    row[age_group] <- round2(as.numeric(row[(age_group - 1)]) + (1/5) * as.numeric(row[(age_group + 4)]) - as.numeric(row[(age_group - 1)]),-2) 
                } else if (age_group + 5 <= max(age_cols) && !is.na(row[age_group + 5])) {
                    row[age_group] <- round2(as.numeric(row[(age_group - 1)]) + (1/6) * as.numeric(row[(age_group + 5)]) - as.numeric(row[(age_group - 1)]),-2) 
                } else {
                    warning("To much missing information to impute ", row["field"], row["level_code"], age_group)
                    row[age_group] <- 0
                }
            }

        }
      row["did_impute"] <- TRUE
    } else {
      row["did_impute"] <- FALSE
    }
  
    row["safe_values"] <- (as.numeric(row["missing_count"]) == 0 || row["did_impute"] == TRUE) && as.numeric(row["people_count"]) > 20
    
    return(row)
}
annual_income_with_imputation <- apply(ready_to_impute, 1, impute_value) %>% t %>% as_tibble %>% 
    mutate_at(vars('age_15_19', 'age_20_24', 'age_25_29', 'age_30_34', 'age_35_39', 'age_40_44', 'age_45_49', 'age_50_54', 'age_55_59', 'age_60_64', 'age_65_'), funs(as.integer))
head(annual_income_with_imputation)
```

```{r}
dbWriteTable(go_analysis_con, Id(schema = "wide", table = "cen_annual_income_with_imputation"), annual_income_with_imputation, overwrite = TRUE)
```

# Construct yearly income estimates based on five year age bands

```{r}
age_lookup <- tibble(age_names = colnames(annual_income_with_imputation)[6:15], midpoint = seq(17, 62, by = 5))
age_lookup
```

```{r}
annual_income_with_imputation %>% 
    filter(safe_values == TRUE, field_code == '0609', level_code == '10')
```

```{r}
age_value_matrix <- annual_income_with_imputation %>% 
    filter(safe_values == TRUE) %>%
    select(-c(missing_count, can_impute, did_impute, safe_values, people_count, age_15_19_impute_income)) %>% apply(1, function(row) {
      for(age_index in 1:(nrow(age_lookup) - 1)) {
          lower_age_band <- age_lookup[age_index,]$age_names
          upper_age_band <- age_lookup[age_index + 1,]$age_names
  
          first_age <- age_lookup[age_index,]$midpoint
          row[as.character(first_age)] <- as.numeric(row[lower_age_band])
          # as.character(i)
          for(i in (first_age+1):(first_age+4)) {
  
  
              row[as.character(i)] <- ((as.numeric(row[upper_age_band]) - as.numeric(row[lower_age_band])) / 5) + as.numeric(row[as.character(i - 1)])
          }
      }
      
      lower_age_band <- "age_60_64"
      upper_age_band <- "age_65_"
      first_age <- 62
      row[as.character(first_age)] <- as.numeric(row[lower_age_band])
      # as.character(i)
      for(i in (first_age+1):(first_age+2)) {
          row[as.character(i)] <- ((as.numeric(row[upper_age_band]) - as.numeric(row[lower_age_band])) / 3) + as.numeric(row[as.character(i - 1)])
      }
      row[as.character(65)] <- as.numeric(row[upper_age_band])
      #i <- i + 1
      #row[as.character(i)] <- ((as.numeric(row[upper_age_band]) - as.numeric(row[lower_age_band])) / 5) + as.numeric(row[as.character(i - 1)])
      return(row)
  })
#head(age_value_matrix)
```


## Check that the columns will be named as expected
```{r}
colnames(age_value_matrix %>% t)
```

## Convert to long format and tibble
```{r}
#Data is in wide format. Age brackets are replaced with individual ages 17-65
yearly_gross_income_long <- age_value_matrix %>% t %>% as_tibble %>% 
    select(-c(age_lookup$age_names, "age_65_")) %>%
    pivot_longer(cols = -c(employed_code, level_code, level, field_code, field), names_to = "age", values_to = "gross_income") %>%
    mutate(gross_income = as.integer(gross_income), age = as.integer(age))
head(yearly_gross_income_long)
```

```{r}
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "cen_yearly_gross_income"), yearly_gross_income_long, overwrite = TRUE)
```