---
title: "Age Standard Income"
output:
  html_document:
    df_print: paged
---

Constuct age standard income. I don't know why doing this is being done, not really.

The *age bands* are *added* together using weights from the working population age distribution. I don't know how this should be interpreted, but it was done last time, so I *assume* the approach is valid.

# Setup

## Load libraries and scripts
```{r}
library(tidyverse)
library(dbplyr)
library(DBI)

source("../scripts/unz_common.R")
source("../scripts/impute_income.R")
source("../scripts/compute_age_standard_income.R")

source("../scripts/cross_tabulation.R")
```

## Connect to analysis database

```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

## Read dataset
```{r}
data <- dbGetQuery(go_analysis_con, "SELECT *
  FROM [lng].[cen_annual_avg_income]
  where set_standard = 1")
head(data)
```

## Read imputed dataset
```{r}
annual_income_with_imputation <- dbReadTable(go_analysis_con, Id(schema = "wide", table = "cen_annual_income_with_imputation"))
head(annual_income_with_imputation)
```


## convert imputed income dataset to long by age
```{r}
income_long <- annual_income_with_imputation %>% filter(can_impute == TRUE) %>%
    select(-c(Total, missing_count, can_impute, age_15_19_impute_income)) %>%
    pivot_longer(names_to = 'age', values_to = 'mean_income',
        cols = c(age_15_19, age_20_24, 
                 age_25_29, age_30_34, 
                 age_35_39, age_40_44, 
                 age_45_49, age_50_54,
                 age_55_59, age_60_64, age_65_
                )
    ) %>%
    filter(!is.na(mean_income)) %>%
    mutate(employed_code = as.numeric(employed_code))
head(income_long)
```

## define rename from column naming back to names with spaces
Used to change the text to human readable for a column, also needed to make some joins work. 
```{r}

convert_to_unsafe_age <- c("age_15_19" = "15-19 years", "age_20_24" = "20-24 years",    
"age_25_29" = "25-29 years", "age_30_34" = "30-34 years", "age_35_39" = "35-39 years",     
"age_40_44" = "40-44 years", "age_45_49" = "45-49 years", "age_50_54" = "50-54 years",   
"age_55_59" = "55-59 years", "age_60_64" = "60-64 years", "age_65_" = "65 years and over")


convert_to_unsafe_age       
```

## Get the age distribution
```{r}
get_income_age_distribution <- function(.data) {
    age_distribution <- .data %>% group_by(employed_code, age) %>% 
        summarise(people_count = sum(people_count, na.rm = TRUE)) %>%
        group_by(employed_code) %>%
        mutate(proportion = people_count/sum(people_count, na.rm = TRUE))
    return (age_distribution)
}
age_distribution <- data %>% select(-c(people_count)) %>% 
    rename(people_count = people_count2) %>% 
    filter(age != 'Total') %>% get_income_age_distribution()
age_distribution
```

## Add age bracket %'s and new age names to long income data
```{r}
income_long %>% mutate(
    age = convert_to_unsafe_age[age]
) %>% left_join(age_distribution, by=c("employed_code", "age")) %>%
group_by(employed_code, level_code, level, field_code, field)
```

## get people counts
```{r}
all_ft_people_counts <- data %>% select(-c(people_count)) %>% 
    rename(people_count = people_count2) %>% 
    filter(age != 'Total') %>%
    group_by(employed_code, level_code, level, field_code, field) %>%
    summarise(people_count = sum(people_count, na.rm = TRUE)) %>%
    filter(people_count > 0)
head(all_ft_people_counts)
```

# age standardise
```{r}
age_standard_income <- income_long %>% 
    mutate(
        age = convert_to_unsafe_age[age]
    ) %>% left_join(age_distribution, by=c("employed_code", "age")) %>%
    group_by(employed_code, level_code, level, field_code, field) %>%
    summarise(age_sd_income = round2(weighted.mean(mean_income, proportion), -2)) %>%
    left_join(all_ft_people_counts) %>%
    mutate(Field_Text = paste(field_code, field)) %>%
    filter(people_count >= 20)
head(age_standard_income)
```

# Output data
```{r}
dbWriteTable(go_analysis_con, 
             Id(schema = "lng", table = "cen_inc_age_sd_annual"), 
             age_standard_income %>% ungroup() %>% select(employed_code, Field_Text, level_code, age_sd_income, people_count), 
             overwrite = TRUE)
```

## output standardised income
```{r}
tabulated_income_a <- age_standard_income %>% pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = "age_sd_income")
print(tabulated_income_a)
dbWriteTable(go_analysis_con, Id(schema = "tab", table = "cen_inc_age_sd_annual"), tabulated_income_a, overwrite = TRUE)
```

## output people counts
```{r}
tabulated_counts <- age_standard_income %>% pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = "people_count")
print(tabulated_counts)
dbWriteTable(go_analysis_con, Id(schema = "tab", table = "cen_inc_age_sd_annual_counts"), tabulated_counts, overwrite = TRUE)
```

