---
title: "Lifetime income analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


The purpose of this is to produce a sound income

This produces the following outputs
- 1a simple lifetime earnings
- 1b offset lifetime earnings
- 2a lifetime earnings above school leaver
- 3a discounted lifetime earnings
- 3b discounted lifetime earnings above a school leaver
- 4 better off than a school leaver
- 5a tax paid over lifetime

```{r}
library(tidyverse)
library(readxl)
library(openxlsx)
library(data.table)
library(dbplyr)
library(DBI)
source("../scripts/unz_common.R")
source("../scripts/impute_income.R")
source("../scripts/compute_age_standard_income.R")
source("../scripts/cross_tabulation.R")
```

## Defining functions to output data 
One for all qual levels, one for levels 4 to 10
```{r}
# for outputting a table into for all study levels
# requires: level_code, Field_Text
# accepts long format
# chainable
output_field_levels_all <- function(.data, con, value_var, table_name) {
    tabulated_income_b <- .data %>% pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = value_var)
    print(tabulated_income_b)
    dbWriteTable(con, Id(schema = "tab", table = table_name), tabulated_income_b, overwrite = TRUE)
    return(.data)
}

# Levels 4-10
output_field_levels_l4_l10 <- function(.data, con, value_var, table_name) {
    tabulated_income_b <- .data %>% pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = value_var) %>%
        select(Field_Text, `lvl 04`, `lvl 05`, `lvl 07`, `lvl 08`, `lvl 09`, `lvl 10`)
    print(tabulated_income_b)
    dbWriteTable(con, Id(schema = "tab", table = table_name), tabulated_income_b, overwrite = TRUE)
    return(.data)
}
```

## Connect to analysis database
```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

## Load Data
Getting combined full-time and part-time data (employed_code)
```{r}
#employed_code = '2'
yearly_gross_income_long <- dbGetQuery(go_analysis_con, "SELECT *
  FROM [lng].[cen_yearly_gross_income]")
head(yearly_gross_income_long)
```

### Load number of people

The yearly income dataset does not contain people counts, so it is being obtained from a different table
```{r}
people_counts <- dbGetQuery(go_analysis_con, 
      "SELECT
      employed_code, [level_code]
      ,[field], field_code, age
      ,[people_count2]
  FROM [GO_Analysis_Data].[lng].[cen_annual_avg_income]
  where people_count2 > 0 and set_standard = 1")

total_people_counts <- people_counts %>% group_by(employed_code, level_code, field, field_code) %>% summarise(people_count2 = sum(people_count2)) %>%
rename(people_count_all_ages = people_count2) %>% 
  mutate(
    Field_Text = paste(field_code, field),
    employed_code = as.character(employed_code)
  ) %>%
    select(-c(field, field_code))
head(total_people_counts)
```

## Check what level codes are in the dataset
```{r}
unique(yearly_gross_income_long[c("level", "level_code")])
```


# Calculate a simple lifetime income

Simple in that it does not remove fees from the total income

```{r}
# simple gross lifetime incomes
simplistic_lifetime_earnings <- yearly_gross_income_long %>% filter(age >= 18, level != "Overseas Secondary School Qualification") %>%
    group_by(employed_code, field, level_code) %>%
    summarise(simple_lifetime_income = sum(gross_income)) %>% 
    append_narrow_field_of_study_code(field_var = "field") %>% 
    mutate(Field_Text = paste(field_code, field))
head(simplistic_lifetime_earnings)
simplistic_lifetime_earnings %>% 
    output_field_levels_all(con = go_analysis_con, value_var = "simple_lifetime_income", table_name = "cen_lifetime_earnings_a")
```



# ******************************************************* 
# This is about where everything gets complex 
# *******************************************************



# Get fee estimates

```{r}
fees_and_funding <- read.csv("../program_costs/fees_durations_SAC.csv") %>% mutate(
  level_code = formatC(level_code, width = 2, flag = "0"),
  field_code = formatC(field_code, width = 4, flag = "0")
  
)
head(fees_and_funding)
```


```{r}
income_with_fees <- yearly_gross_income_long %>%
    left_join(fees_and_funding, by=c("field_code", "level_code"))
head(income_with_fees)
```



```{r}
tax_calculated <- income_with_fees %>% group_by(employed_code, field, level) %>% arrange(age) %>% 
    mutate( #Offset income = income, unless younger than minimum qual completion age
        offset_income = ifelse(is.na(duration_whole_yr), gross_income, ifelse(age < 17 + duration_whole_yr, 0, gross_income)),
         #Loan payments are 12% of all income over $19760/year
        offset_income_loanable = ifelse(offset_income < 19760, 0, (offset_income-19760)),
        loan_repayments = ifelse(
          offset_income < 19760, 0, ifelse(
            #If loan payments up till now have covered loan, repayment = 0
            (cumsum(offset_income_loanable*0.12) - (offset_income_loanable*0.12)) >= loan_at_graduation, 0, ifelse(
              #Otherwise, if this year's payment will be more than enough, repayment = remaining amount
              cumsum(offset_income_loanable*0.12) >= loan_at_graduation, 
              (loan_at_graduation - (cumsum(offset_income_loanable*0.12)-(offset_income_loanable*0.12))), 
              #Otherwise, repayment = 12%
              0.12*offset_income_loanable
            )
          )
        )
        
    ) %>% 
    mutate( # Calculating tax paid each year using marginal tax rates
        current_loan = loan_at_graduation - cumsum(loan_repayments),
        tax = ifelse(
            offset_income <= 14000, 
            offset_income * 0.105,
            ifelse(
                offset_income <= 28000, 
                1470 + (offset_income - 14000) * 0.175,
                ifelse(
                    offset_income <= 70000,
                    3920 + (offset_income - 28000) * 0.30,
                    16520 + (offset_income - 70000) * 0.33
                )
            )
        )
    )
head(tax_calculated)
```

Calculating baseline tax and income for each year for a school leaver with no post-school qualification
```{r}
baseline_income <- tax_calculated %>% 
    filter(field_code == "0000", level_code == "00") %>%
    ungroup() %>%
    select(employed_code, age, gross_income, tax) %>%
    rename(baseline_income = gross_income, baseline_tax = tax)
head(baseline_income)
```

Calculating difference in earnings and tax for each qual from baseline
```{r}
components_calculated <- tax_calculated %>% left_join (baseline_income, by=c("employed_code", "age")) %>%
    mutate(
        above_school_leaver = offset_income - baseline_income - loan_repayments,
        tax_paid_above_school_leaver = tax - baseline_tax
    )

head(components_calculated)
```



# Calculating Net Present Value (NPV)

NPV = COST OF INVESTMENT - SUM(PRESENT VALUE RETURN FOR EACH TIME PERIOD)
PRESENT VALUE RETURN FOR EACH TIME PERIOD = VALUE * (1/(1+inflation rate))^no. periods

or, discounted_income_tax_paid_above_school_leaver_m_sac = tax_paid_above_school_leaver_m_sac * (v ^ t),
```{r}
i <- 0.03 #Discount rate or return that could be made from alternative investment (in this case, not paying SAC funding) = INFLATION RATE = 3%
v <- 1 / (1+i)
print(c(i,v))
npv_calculated <- components_calculated %>% 
    mutate(
        t = (age - 17),
        discounted_income = offset_income * (v ^ t),
        npv = cumsum(discounted_income),
        discounted_compared_school_leaver = above_school_leaver  * (v ^ t),
        npv_over_school_leaver = cumsum(discounted_compared_school_leaver),
        
        # SAC funding:
        sac_fund_for_yr_temp = sac_funding * 1,
        sac_fund_for_yr = 
            ifelse(
                cumsum(sac_fund_for_yr_temp) > total_funding, 
                ifelse(
                    cumsum(sac_fund_for_yr_temp) - sac_fund_for_yr_temp > total_funding,
                    0,
                    total_funding - (cumsum(sac_fund_for_yr_temp) - sac_fund_for_yr_temp)
                ),
                sac_fund_for_yr_temp
            ),
        tax_paid_above_school_leaver_m_sac = tax_paid_above_school_leaver - sac_fund_for_yr,
        
        discounted_income_tax_paid_above_school_leaver_m_sac = tax_paid_above_school_leaver_m_sac * (v ^ t),
        npv_tax_over_school_leaver = cumsum(discounted_income_tax_paid_above_school_leaver_m_sac)
    )
head(npv_calculated)
```


# Lifetime income
```{r}
summarised_lifetime_income <- npv_calculated %>% mutate(Field_Text = paste(field_code, field)) %>% 
    group_by(employed_code, Field_Text, level_code) %>%
    summarise(
        lifetime_above_school_leaver = sum(above_school_leaver), 
        discounted_lifetime_income = sum(discounted_income), 
        discounted_lifetime_above_school_leaver = sum(discounted_compared_school_leaver), 
        offset_lifetime_income = sum(offset_income),
        lifetime_tax = sum(tax),
        tax_above_school_leaver = sum(tax_paid_above_school_leaver),
        
        npv_tax_over_school_leaver = sum(discounted_income_tax_paid_above_school_leaver_m_sac)
    ) %>% left_join (total_people_counts) %>% filter (people_count_all_ages > 25)
head(summarised_lifetime_income)
```


```{r}
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "cen_lifetime_income"), 
             summarised_lifetime_income %>% filter(
                ! (is.na(lifetime_above_school_leaver) & is.na(discounted_lifetime_income) & is.na(discounted_lifetime_above_school_leaver) & is.na(offset_lifetime_income) & is.na(lifetime_tax) & is.na(tax_above_school_leaver))
             ), 
             overwrite = TRUE)
```

```{r}
tabulated_income_b <- summarised_lifetime_income %>% pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = "lifetime_above_school_leaver")
print(tabulated_income_b)
dbWriteTable(go_analysis_con, Id(schema = "tab", table = "cen_lifetime_earnings_above_school_leaver"), tabulated_income_b, overwrite = TRUE)
```

```{r}
summarised_lifetime_income %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "lifetime_above_school_leaver", "cen_lifetime_earnings_above_school_leaver") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "discounted_lifetime_income", "cen_lifetime_discounted_earnings") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "discounted_lifetime_above_school_leaver", "cen_lifetime_discounted_earnings_above_school_leaver") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "offset_lifetime_income", "cen_lifetime_offset") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "lifetime_tax", "cen_lifetime_tax") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "tax_above_school_leaver", "cen_tax_above_school_leaver") %>%
    output_field_levels_l4_l10(con = go_analysis_con, value_var = "npv_tax_over_school_leaver", "cen_npv_tax_over_school_leaver") %>%
    head()
```


Finding age at which graduates are better off financially than school-leavers, by finding the year at which npv_over_school_leaver becomes positive
```{r}
age_when_better_off <- npv_calculated %>% mutate(
    better_off = ifelse(lag(npv_over_school_leaver) <= 0 & npv_over_school_leaver > 0, TRUE, FALSE),
    Field_Text = paste(field_code, field)
) %>% group_by(employed_code, Field_Text, level_code) %>%
arrange(Field_Text, level_code, desc(better_off), age) %>%
mutate(
    fl_index = row_number()
) %>% filter(!is.na(better_off), fl_index == 1) %>% ungroup() %>%
    select(employed_code, Field_Text, level_code, age, better_off) %>%
    mutate(age = ifelse(!better_off, 99, age)) %>%
    arrange(Field_Text, level_code)

head(age_when_better_off)
```

```{r}
tabulated_age_better_off <- age_when_better_off %>% 
    pivot_wider(id_cols = c("employed_code", "level_code", "Field_Text"), names_from = level_code,names_prefix = "lvl ", values_from = "age") %>%
    select(employed_code, Field_Text, `lvl 04`, `lvl 05`, `lvl 07`, `lvl 08`, `lvl 09`, `lvl 10`)

print(tabulated_age_better_off)
dbWriteTable(go_analysis_con, Id(schema = "tab", table = "cen_age_better_off"), tabulated_age_better_off, overwrite = TRUE)
```



```{r}
# do some checks
npv_calculated %>% filter(field_code == "0201", level_code == "07") %>%
    select(age, gross_income)
```

# Output age when better off in long format
```{r}
dbWriteTable(go_analysis_con, Id(schema = "lng", table = "age_better_off"), age_when_better_off, overwrite = TRUE)
```

Potential further analysis: look at age at which government is better off (i.e. SAC funding has paid off in increased tax revenue from higher-earning graduates)