---
title: "Calculate Typical Study Level"
output: html_notebook
---

inline code? `r "printed"`

```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(readxl)
```


# Read occupation file
```{r}
occupation_studylevel <- read_excel("../input_data/t3 - Occupation by field of study - Checked.xlsx", sheet="Occupation by level totals", skip = 4)
head(occupation_studylevel)
```

## compute occupation counts

this might not be using the occupation totals dataset to exclude not disclosed.
```{r}
occupation_counts <- occupation_studylevel %>% filter(
    !i_standard_highest_qual_code %in% c(97, 99)
) %>% group_by(i_occupation_code) %>%
    summarise(occupation_people_count = sum(people_count))
head(occupation_counts)
```

## pivot occupation by study level
```{r}
occupation_studylevel_wide <- occupation_studylevel %>% pivot_wider(names_from = i_standard_highest_qual_code, names_prefix = "lvl_", values_from = people_count)
head(occupation_studylevel_wide)
```

## calculate how many are at different study level groups
```{r}
result <- occupation_studylevel_wide %>%
        group_by(i_occupation_code) %>%
        mutate(school = sum(lvl_0, lvl_1, lvl_2, lvl_3, lvl_11, na.rm = TRUE), 
               cert_dip = sum(lvl_5, na.rm = TRUE),
               degree_or_higher = sum(lvl_7, lvl_8, lvl_9, lvl_10, na.rm = TRUE), 
               tertiary = sum(lvl_4, lvl_5, lvl_7, lvl_8, lvl_9, lvl_10, na.rm = TRUE)
              ) %>% ungroup()
head(result)
```

## merge with occupation counts and check proportions for the study level groups
```{r}
standard_threshold <- 0.85
conditions_applied <- result %>% inner_join(occupation_counts) %>%
    mutate(
        school_yes = school / occupation_people_count > standard_threshold,
        cert_dip_yes = cert_dip / occupation_people_count > standard_threshold,
        degree_or_higher_yes = degree_or_higher / occupation_people_count > standard_threshold,
        tertiary_yes = tertiary / occupation_people_count > 0.90
    )
head(conditions_applied)
```

## Using a prioritied method asign a typical study level classification number to each occupation

1 = school
2 = certificates and diplomas
3 = mixed
4 = tertiary
5 = degree
```{r}
typical_study_levels <- conditions_applied %>%
    mutate(
        typical_study = ifelse(
            school_yes,
            1,
            ifelse(
                cert_dip_yes,
                2,
                ifelse(
                    degree_or_higher_yes,
                    5,
                    ifelse(
                        tertiary_yes,
                        4,
                        3
                    )
                )
            )
        )
    ) %>% select(i_occupation_code, typical_study)
head(typical_study_levels)
```

## Connect to analysis database

```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

## Write typical study levels table to database
```{r}
dbWriteTable(go_analysis_con, name = "typical_study_levels",  
                typical_study_levels
             , overwrite = TRUE)
```