---
title: "Calculate Typical Study Level"
output:
  html_document:
    df_print: paged
---

inline code? `r "test"`

The objective of this notebook is to calculate the typical study level of individuals in different occupations.

This is done by calculating how many individuals are at different levels, making that a proportion and adding them together to see if any of the groups are met.

The groups are (in this order):

* school (level 0 - 3)
* certificates and diplomas (level 4 - 6) - CURRENTLY APPLIES TO NO OCCUPATIONS
* degree (level 7+)
* tertiary (level 4+)
* mixed (if none of the above selected)


# Setup
## Libs
```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(readxl)
```


## Read occupation file
```{r}
occupation_studylevel <- read_excel("../input_data/t3 - Occupation by field of study - Checked.xlsx", sheet="Occupation by level totals", skip = 4)
head(occupation_studylevel)
```

## compute occupation counts

Excluding not disclosed and unknown qualification levels.
```{r}
occupation_counts <- occupation_studylevel %>% filter(
    !i_standard_highest_qual_code %in% c(97, 99)
) %>% group_by(i_occupation_code) %>%
    summarise(occupation_people_count = sum(people_count))
head(occupation_counts)
```

## pivot occupation by study level (converting to wide format)
```{r}
occupation_studylevel_wide <- occupation_studylevel %>% pivot_wider(names_from = i_standard_highest_qual_code, names_prefix = "lvl_", values_from = people_count)
head(occupation_studylevel_wide)
```

## calculate how many are at different study level groups
```{r}
result <- occupation_studylevel_wide %>%
        group_by(i_occupation_code) %>%
        mutate(school = sum(lvl_0, lvl_1, lvl_2, lvl_3, lvl_11, na.rm = TRUE), 
               cert_dip = sum(lvl_5, na.rm = TRUE),
               degree_or_higher = sum(lvl_7, lvl_8, lvl_9, lvl_10, na.rm = TRUE), 
               tertiary = sum(lvl_4, lvl_5, lvl_7, lvl_8, lvl_9, lvl_10, na.rm = TRUE)
              ) %>% ungroup()
head(result)
```

## merge with occupation counts and check proportions for the study level groups
True or false indicators for whether the occupation is dominated by people with a certain (grouped) qualification level, using 85% and 90% thresholds
```{r}
standard_threshold <- 0.85
conditions_applied <- result %>% inner_join(occupation_counts) %>%
    mutate(
        school_yes = school / occupation_people_count > standard_threshold,
        cert_dip_yes = cert_dip / occupation_people_count > standard_threshold,
        degree_or_higher_yes = degree_or_higher / occupation_people_count > standard_threshold,
        tertiary_yes = tertiary / occupation_people_count > 0.90
    )
head(conditions_applied)
```

NOTE: Currently no qualifications have cert_dips as dominant level of study

## Using a prioritied method asign a typical study level classification number to each occupation

1 = school
2 = certificates and diplomas
3 = mixed
4 = tertiary
5 = degree
```{r}
typical_study_levels <- conditions_applied %>%
    mutate(
        typical_study = ifelse(
            school_yes,
            1,
            ifelse(
                cert_dip_yes,
                2,
                ifelse(
                    degree_or_higher_yes,
                    5,
                    ifelse(
                        tertiary_yes,
                        4,
                        3
                    )
                )
            )
        )
    ) %>% select(i_occupation_code, typical_study)
head(typical_study_levels)
```

## Connect to analysis database

```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

## Write typical study levels table to database
```{r}
dbWriteTable(go_analysis_con, name = "typical_study_levels",  
                typical_study_levels
             , overwrite = TRUE)
```