---
title: "Generate Graduate ROI Executive"
output:
  html_document:
    df_print: paged
---

The main goal of this notebook is to create a formatted excel workbook containing a couple of sheets, relating to most common occupations and qualifications, used by the executive director (Chris) to answer ad-hoc queries.

```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(readxl)
library(openxlsx)
library(foreach)
source("../scripts/unz_common.R")
```

## Connect to analysis database

```{r}
go_analysis_con <- connect_to_GO_analysis_database()
go_analysis_con
```

# Read input file
```{r}
grad_outcomes_datacube <- read.csv("../input_data/occupations_by_levels_narrow_2018.csv")
head(grad_outcomes_datacube)
```

#Adjusting format of columns, bringing in text variables for occupation and narrow field of study
```{r}
grad_cube2 <- grad_outcomes_datacube %>% mutate(i_occupation_code = as.character(i_occupation_code)) %>%
    append_ANZSCO_L5_txt(code_var = "i_occupation_code", ANZSCO_var = "occupation_text") %>%
#    mutate(narrow_field_code = formatC(narrow_field_code, width = 4, format = "d", flag = "0")) %>%
    append_narrow_field_of_study_code(code_var = "narrow_field_code", field_var = "narrow_field_text") %>%
    append_study_level_code(level_var = "i_standard_highest_qual_code", code_var = "level_code") %>%
    rename(i_standard_highest_qual_text = i_standard_highest_qual_code, i_standard_highest_qual_code = level_code) %>%
    filter(!i_standard_highest_qual_code %in% c(97, 99, NA)) %>% #Remove people with highest qualifications Not Elsewhere Included
    mutate(
        i_standard_highest_qual_code =  formatC(i_standard_highest_qual_code, width = 2, format = "d", flag = "0"),
        i_standard_highest_qual_code = ifelse(i_standard_highest_qual_code == '11', 'OS', i_standard_highest_qual_code)
    ) %>%  filter(people_count!=0)
grad_cube2$narrow_field_code[is.na(grad_cube2$narrow_field_code)] <- "0000"
grad_cube2$narrow_field_code[grad_cube2$narrow_field_text=='Field of Study Not Given'] <- "3333"
grad_cube2$narrow_field_text[grad_cube2$narrow_field_text=='No Qualification'] <- "No Post-School Qualification"
head(grad_cube2)
```


# get top 30 occupations
```{r}
top_occupations_1 <- grad_cube2 %>% mutate(
    occupation_id = paste(i_occupation_code, occupation_text),
    occupation = paste0(occupation_id, ' (', people_count, ')')
) %>% group_by(
    narrow_field_code, narrow_field_text, i_standard_highest_qual_code
) %>% arrange(narrow_field_code, narrow_field_text, i_standard_highest_qual_code) %>% 
    arrange(desc(people_count), .by_group = TRUE) %>%
    mutate(
        occupationRank = 1:n(),
        rankText = nth(v = occupationRank, first_value =  "top occupation", second_value = '2nd occupation'),
        field_id = paste0(narrow_field_code, ' ' , narrow_field_text, ' [lvl-', i_standard_highest_qual_code, ']')
    ) %>%
filter(occupationRank <= 30) %>% select(
    field_id, narrow_field_code, narrow_field_text, i_standard_highest_qual_code, occupationRank, rankText, occupation
) 
head(top_occupations_1)
```

Converting to wide form, with 30 columns for each top occupation for each qual
```{r}
top_occupations_2 <- top_occupations_1 %>% ungroup() %>% 
    select(-c(occupationRank, narrow_field_text)) %>% 
    pivot_wider(id_cols = c("field_id", "narrow_field_code", "i_standard_highest_qual_code"), values_from = occupation, names_from = "rankText")
head(top_occupations_2)
```

#Getting number of people for each field of study, with correct format
```{r}
field_of_study_totals <- read.csv("../input_data/field_counts_2018.csv") %>%
append_narrow_field_of_study_code(code_var = "narrow_field_code", field_var = "narrow_field_text") %>%
  select(-narrow_field_text)
head(field_of_study_totals)
```

## Read field of study income 
Dataset is already limited to full-time (employment code 1) employees
```{r}
consolidated_income <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "consolidated_income")) %>%
    mutate(age_when_better_off = ifelse(ever_better_off, age_when_better_off, "NEVER"))
head(consolidated_income)
```

Need to do conversions of id columns to match
Then rename columns as desired (possibly use spaces?)
```{r}
income_table <- consolidated_income %>% mutate(narrow_field_code = substr(Field_Text,1,4), BLANK_A = NA) %>%
    rename(
        i_standard_highest_qual_code = level_code,
        `Avg Fulltime Annual Income 30-39 Yr Olds` = ft_30_39_annual_income,
        `Total Fulltime Employed 30-39 Yr Olds` = ft_30_39_people_count,
        `Avg Annual Income for all 30-39 Yr Olds` = all_30_39_annual_income,
        `Total Employed all 30-39 Yr Olds` = all_30_39_people_count,
        `Avg Annual Income for 20-65+ Yr Olds Fulltime Only` = ft_sd_annual_income,
        `Avg Annual Income for 20-65+ Yr Olds Fulltime & Part Time` = all_sd_annual_income,
        `Total Employed 20-65+ Yr Olds Fulltime Only` = ft_people_count,
        `Total Employed 20-65+ Yr Olds Fulltime & Part Time` = all_people_count,
        `Additional lifetime earnings above a school leaver 20-65+` = lifetime_above_school_leaver,
        `Whole of life NPV for grads at Lvl 4 & above in full time employment 20-65+` = discounted_lifetime_above_school_leaver,
        `Age when better off as a graduate (NPV) Lvl 4+ only` = age_when_better_off,
        `NPV of income tax paid to Govt above school leaver` = npv_tax_over_school_leaver
    ) %>% select(-c(
        discounted_lifetime_income,
        offset_lifetime_income,
        lifetime_tax,
        tax_above_school_leaver,
        people_count_all_ages,
        ever_better_off
    )) 

head(income_table)
```

Getting top occupations
```{r}
top_occupations_3 <- income_table %>% right_join(top_occupations_2, by=c("narrow_field_code", "i_standard_highest_qual_code")) %>% 

    select(-Field_Text) %>% select(field_id, narrow_field_code, i_standard_highest_qual_code, everything()) %>% select(-narrow_field_code)
head(top_occupations_3)
```

Saving table to database
```{r}
dbWriteTable(go_analysis_con, name = "top_occupations", top_occupations_3, overwrite = TRUE)
```

Transposing table
```{r}
top_occupations_4 <- top_occupations_3 %>% select(-i_standard_highest_qual_code) %>% t %>% as.data.frame(stringsAsFactors = FALSE) 
head(top_occupations_4)
```

# Now getting most common fields of study by occupation
```{r}
top_fields_1 <- grad_cube2 %>% mutate(
    field_id = paste('lvl', i_standard_highest_qual_code, narrow_field_text),
    field = paste0(field_id, ' (', people_count, ')')
) %>% group_by(
    i_occupation_code, occupation_text
) %>% arrange(i_occupation_code, occupation_text) %>% 
    arrange(desc(people_count), .by_group = TRUE) %>%
    mutate(
        fieldRank = 1:n(),
        rankText = nth(v = fieldRank, first_value =  "top field of study", second_value = '2nd field of study'),
        occupation_id = paste(occupation_text)
    ) %>%
filter(fieldRank <= 30) %>% ungroup() %>%
select(occupation_id, i_occupation_code, occupation_text, field, fieldRank, rankText)

head(top_fields_1)
```

```{r}
top_fields_2 <- top_fields_1 %>% ungroup() %>% select(-fieldRank) %>% pivot_wider(values_from = field, names_from = "rankText")
head(top_fields_2)
```

```{r}
 typical_study_levels <- dbReadTable(go_analysis_con, name = "typical_study_levels") %>% mutate(i_occupation_code = as.character(i_occupation_code))
typical_study_levels$typical_study <- factor(typical_study_levels$typical_study, levels = 1:5, labels = c("School", "Cert/Dipl", "Mixed", "Tertiary","Degree"))
head(typical_study_levels)
```

```{r}
occupation_totals <- read.csv("../input_data/occup_counts_2018.csv")
occupation_totals$i_occupation_code <- as.character(occupation_totals$i_occupation_code)
head(occupation_totals)
```


```{r}
top_fields_3 <- top_fields_2 %>% 
    left_join(typical_study_levels) %>% 
    left_join(occupation_totals) %>%
    
    select(occupation_id, i_occupation_code, occupation_text, people_count, typical_study, everything()) %>% select(-c(occupation_text))
head(top_fields_3)
```

```{r}
top_fields_4 <- top_fields_3 %>% t %>% as.data.frame(stringsAsFactors = FALSE)
head(top_fields_4)
```

# Build cross reference tables
```{r}
degree_keys <- read_excel("../metadata/degree_key_lookup.xlsx") %>% select(-occupation_text) %>% rename(i_occupation_code = anzsco_code)
head(degree_keys)
```



```{r}
occupation_cross_ref_table <- grad_cube2 %>% mutate(
    field_id = paste0(narrow_field_code, ' ', narrow_field_text, ' [lvl-', i_standard_highest_qual_code, ']')
) %>%
    pivot_wider(id_cols = c(occupation_text, i_occupation_code), names_from = field_id, values_from = people_count) %>%
    arrange(occupation_text) %>% left_join(degree_keys) %>% select(-i_occupation_code)
head(occupation_cross_ref_table)
```

## Build field of study at top sheet
```{r}
occupation_cross_ref_table_1 <- grad_cube2 %>% mutate(
    field_id = paste0(narrow_field_code, ' ', narrow_field_text, ' [lvl-', i_standard_highest_qual_code, ']')
) %>% arrange(occupation_text) %>%
    pivot_wider(id_cols = field_id, names_from = occupation_text, values_from = people_count) %>%
    arrange(field_id) %>%
    add_column( # columns to add space between the grad cube and top occupations
        BLANK1 = NA,
        BLANK2 = NA,
        BLANK3 = NA,
        BLANK4 = NA,
        BLANK5 = NA,
        BLANK6 = NA,
        BLANK7 = NA,
        BLANK8 = NA,
        BLANK9 = NA,
        BLANK10 = NA
    ) %>%
    full_join(top_occupations_3 %>% rename(`Narrow Field of Study` = field_id) %>% mutate(field_id = `Narrow Field of Study`), by=c("field_id"))
head(occupation_cross_ref_table_1)
```

### Add degree keys
```{r}
occupation_table_5 <- occupation_cross_ref_table_1 %>% 
    select(-i_standard_highest_qual_code) %>% 
    t %>% as.data.frame(stringsAsFactors = FALSE) %>%
    rownames_to_column("id") %>% 
    mutate(
        id = ifelse(startsWith(id, "BLANK"), NA, id)
    ) %>% left_join(degree_keys %>% append_ANZSCO_L5_txt(code_var = "i_occupation_code", ANZSCO_var = "id") %>% select(-i_occupation_code) %>% filter(!is.na(id)), by="id")
occupation_table_5$degree_key[1] <- "Degree Keys"
head(occupation_table_5)
tail(occupation_table_5)
```

## Build occupation at top sheet table

```{r}
field_level_cross_ref_table <- grad_cube2 %>% mutate(
    occupation_id = paste0(occupation_text)
) %>% arrange(occupation_id) %>%
    pivot_wider(id_cols = c(narrow_field_text, i_standard_highest_qual_code, narrow_field_code), names_from = occupation_id, values_from = people_count) %>%
    arrange(narrow_field_code)
head(field_level_cross_ref_table)
```

top_fields_3
```{r}
field_level_cross_ref_table_1a <- grad_cube2 %>% 
    mutate(
        occupation_id = paste0(occupation_text),
        field_id = paste0(narrow_field_code, ' ', narrow_field_text, ' [lvl-', i_standard_highest_qual_code, ']')
    ) %>% arrange(field_id) %>%
    pivot_wider(id_cols = occupation_id, names_from = field_id, values_from = people_count) %>%
    arrange(occupation_id) %>%
    add_column( # columns to add space between the grad cube and top occupations
        BLANK1 = NA,
        BLANK2 = NA,
        BLANK3 = NA,
        BLANK4 = NA,
        BLANK5 = NA,
        BLANK6 = NA,
        BLANK7 = NA,
        BLANK8 = NA,
        BLANK9 = NA,
        BLANK10 = NA
    )
print(ncol(field_level_cross_ref_table_1a))
field_level_cross_ref_table_1 <- field_level_cross_ref_table_1a %>% 
    full_join(
        top_fields_3 %>% 
            rename(`30 Most Common Qualifications for each Job Title for people aged 30-39` = occupation_id) %>%
            mutate(occupation_id = `30 Most Common Qualifications for each Job Title for people aged 30-39`), 
        by="occupation_id"
    )
head(field_level_cross_ref_table_1)
```

```{r}
occupation_top_table <- field_level_cross_ref_table_1 %>% t %>% as.data.frame(stringsAsFactors = FALSE) %>%
    rownames_to_column("id") %>% 
    mutate(
        id = ifelse(startsWith(id, "BLANK"), NA, id)
    )
head(occupation_top_table)
tail(occupation_top_table)
```

# Outputting to Excel workbook
## Lots of Excel formatting follows
```{r}
wb <- createWorkbook('Scott Henwood @ Universities NZ')
# sheet names
occupation_sheet_name <- "Top 30 Qualifications"
field_of_study_sheet_name <- "Top 25 Jobs"
# creating worksheets
addWorksheet(wb, occupation_sheet_name)
addWorksheet(wb, field_of_study_sheet_name)

wb
```



```{r}
primary_occupation_output <- occupation_top_table # top_fields_4
writeData(wb,occupation_sheet_name,primary_occupation_output, rowNames=FALSE, colNames = FALSE)
```

```{r}
apply_top_conditional_formating_study_level = function(standard_args, typical_study_text, colour) {
    conditionalFormatting(
        standard_args$wb, 
        standard_args$sheet, 
        standard_args$colRange, 
        standard_args$row, 
        rule=paste0(standard_args$ts_selector, '="', typical_study_text, '"'), 
        style = openxlsx::createStyle(bgFill = colour)
    )
}


topPartRowCount = ncol(field_level_cross_ref_table_1a)
colRange <- 2:(ncol(primary_occupation_output)+1)
rowRange <- (topPartRowCount+4):(nrow(primary_occupation_output)+3)
#Freeze panes
freezePane(wb, occupation_sheet_name, firstActiveRow = 2,  firstActiveCol = 2)
# apply conditional formating
conditionalFormatting(wb, occupation_sheet_name, colRange, rowRange, rule="INT(MID(INDIRECT(ADDRESS(ROW(),COLUMN())),5,2))>=0", style = openxlsx::createStyle(bgFill = "#F4C724"))
conditionalFormatting(wb, occupation_sheet_name, colRange, rowRange, rule='=MID(INDIRECT(ADDRESS(ROW(),COLUMN())),5,2)="OS"', style = openxlsx::createStyle(bgFill = "#F4C724"))
conditionalFormatting(wb, occupation_sheet_name, colRange, rowRange, rule="INT(MID(INDIRECT(ADDRESS(ROW(),COLUMN())),5,2))>3", style = openxlsx::createStyle(bgFill = "#74B9FF"))
conditionalFormatting(wb, occupation_sheet_name, colRange, rowRange, rule="INT(MID(INDIRECT(ADDRESS(ROW(),COLUMN())),5,2))>=7", style = openxlsx::createStyle(bgFill = "#45CE30"))
#apply static formatting
setColWidths(wb, occupation_sheet_name, cols = colRange, widths = '16')
addStyle(wb, occupation_sheet_name, openxlsx::createStyle(wrapText =TRUE), rows = rowRange, cols = colRange, gridExpand = TRUE)

#apply conditional formating for occupation row
typical_study_row = 3
common_arguments <- list(
    wb = wb,
    sheet = occupation_sheet_name,
    colRange = colRange,
    row = topPartRowCount + 1,
    ts_selector = 'INDIRECT(ADDRESS(ROW()+3,COLUMN()))'
)

typical_study_level_colour_map <- list(
    list(level="School", colour = "#bf8f00"),
    list(level = "Cert/Dipl", colour = "#b4c6e7"),
    list(level = "Mixed", colour = "#ededed"),
    list(level = "Tertiary", colour = "#92d050"),
    list(level = "Degree", colour = "#00ff00")
)

foreach(study_colour = typical_study_level_colour_map) %do%
    apply_top_conditional_formating_study_level(common_arguments, study_colour$level, study_colour$colour)


#apply static formatting
addStyle(wb, occupation_sheet_name, openxlsx::createStyle(wrapText =TRUE, textDecoration = "bold"), rows = topPartRowCount + 1, cols = colRange, gridExpand = TRUE)

common_arguments$row <- topPartRowCount + 4
common_arguments$ts_selector <- 'INDIRECT(ADDRESS(ROW(),COLUMN()))'

foreach(study_colour = typical_study_level_colour_map) %do%
    apply_top_conditional_formating_study_level(common_arguments, study_colour$level, study_colour$colour)

# style first collumn
setColWidths(wb, occupation_sheet_name, cols = 1, widths = '46')
addStyle(wb, occupation_sheet_name, openxlsx::createStyle(textDecoration = "bold", wrapText = TRUE), cols = 1, rows = 1:(nrow(primary_occupation_output)+3), gridExpand = TRUE)
```


```{r}
primary_field_of_study_output <- occupation_table_5 # top_occupations_4
openxlsx::writeData(wb,field_of_study_sheet_name,primary_field_of_study_output, rowNames=FALSE, colNames = FALSE)
```

```{r}
colRange <- 2:(ncol(primary_field_of_study_output)+1)
rowRange <- 2:(nrow(primary_field_of_study_output)+1)

freezePane(wb, field_of_study_sheet_name, firstActiveRow = 2,  firstActiveCol = 2)

addStyle(wb, field_of_study_sheet_name, createStyle(wrapText =TRUE), rows = rowRange, cols = colRange, gridExpand = TRUE)

addStyle(wb, field_of_study_sheet_name, createStyle(wrapText =TRUE, textDecoration = "bold"), rows = 1, cols = colRange, gridExpand = TRUE)

# Set left pane:
setColWidths(wb, field_of_study_sheet_name, cols = colRange, widths = '23')
setColWidths(wb, field_of_study_sheet_name, cols = 1, widths = '30')
addStyle(wb, field_of_study_sheet_name, createStyle(wrapText =TRUE, textDecoration = "bold"), rows = rowRange, cols = 1, gridExpand = TRUE)
```

```{r}
#writeData(wb,field_of_study_cross_name,field_level_cross_ref_table, rowNames=FALSE, colNames = TRUE)
openxlsx::saveWorkbook(wb, '../working_outputs/Graduate ROI Occupation by Field of Study 2018 NEW.xlsx', overwrite = T)
```
