---
title: "Populate GO website database"
output: html_notebook
---


```{r}
library(tidyverse)
library(dbplyr)
library(DBI)
library(readxl)
source("../scripts/unz_common.R")
```

## Database connections
```{r}
go_analysis_con <- DBI::dbConnect(odbc::odbc(),
                      Driver = "SQL Server", 
                      Server = "localhost\\SQLEXPRESS",
                      Database = "GO_Analysis_Data",
                      Trusted_Connection = "True")
go_analysis_con

go_web_con_old <- DBI::dbConnect(odbc::odbc(),
                      Driver = "SQL Server", 
                      Server = "localhost\\SQLEXPRESS",
                      Database = "GradOutcomesNZv4-1",
                      Trusted_Connection = "True")
go_web_con_old

go_web_con <- DBI::dbConnect(odbc::odbc(),
                      Driver = "SQL Server", 
                      Server = "localhost\\SQLEXPRESS",
                      Database = "GradOutcomesNZv5-0",
                      Trusted_Connection = "True")
go_web_con
```

## Populate Broad field of study table
```{r}
dbWriteTable(go_web_con, name = "BroadFieldOfStudy",  
                read_xlsx("../metadata/NZSCED_broad_cen.xlsx", skip = 7) %>%
                    mutate(Code = strtoi(Code, base = 10L)) %>%
                    rename(ClassificationTitle = Descriptor, BroadFieldOfStudyCode = Code)
             , append = TRUE)
```

## Populate Narrow field of study table
Check which sort of total this should be?


```{r}
field_count_totals <- read_xlsx("../input_data/t3b - Occupation by narrow field of study- checked.xlsx", sheet = "narrow field of study totals", skip = 5) %>%
    filter(set_code == 1) %>% select(-set_code) %>%
    mutate(narrow_field_code = formatC(narrow_field_code, width = 4, flag = "0"))
head(field_count_totals)
```

```{r}
field_of_study_table <- field_count_totals %>% append_narrow_field_of_study_txt(code_var = "narrow_field_code", field_var = "ClassificationTitle") %>%
    mutate(
        NarrowFieldOfStudyCode = strtoi(narrow_field_code, base = 10L), 
        BroadFieldOfStudyCode = strtoi(substring(narrow_field_code, 1,2), base = 10L)
    ) %>%
    rename(TotalPeopleCount = people_count) %>%
    select(NarrowFieldOfStudyCode, BroadFieldOfStudyCode, ClassificationTitle, TotalPeopleCount)

head(field_of_study_table)
```

```{r}
dbWriteTable(go_web_con, name = "NarrowFieldOfStudy",  
                field_of_study_table
             , append = TRUE)
```

## Populate Narrow field of study level table
TODO: get data different way
```{r}
# using study requirements numbers from old database.
study_requirements <- dbReadTable(go_web_con_old, "FieldOfStudyStaircasing") %>% 
    select(-c(Id)) %>%
    mutate(NarrowFieldOfStudyCode = formatC(NarrowFieldOfStudyCode, width = 4, flag = '0')) %>%
    rename(StudyRequirementsKey = StaircasingCode)
head(study_requirements)

write.csv(study_requirements, "../metadata/staircasing.csv", row.names = FALSE)
```


```{r}
study_requirements <- read.csv("../metadata/staircasing.csv")
head(study_requirements)
```

```{r}
consolidated_income <- dbReadTable(go_analysis_con, DBI::Id(schema = "lng", table = "consolidated_income"))
head(consolidated_income)
```

```{r}
lifetime_earnings_data <- dbReadTable(go_analysis_con, DBI::Id(schema = "lng", table = "consolidated_income")) %>%
    mutate(NarrowFieldOfStudyCode = strtoi(substring(Field_Text, 1, 4), base = 10L), NZQALevel = strtoi(level_code, base = 10L), EarnedAboveSchoolLeaver = as.integer(lifetime_above_school_leaver)) %>%
    select(NarrowFieldOfStudyCode, NZQALevel, EarnedAboveSchoolLeaver) %>%
    filter(NZQALevel != 11, NZQALevel > 3)
head(lifetime_earnings_data)
```

```{r}
narrow_field_of_study_level <- consolidated_income %>%  mutate(
    NarrowFieldOfStudyCode = strtoi(substring(Field_Text, 1, 4), base = 10L), 
    NZQALevel = strtoi(level_code, base = 10L), 
    EarnedAboveSchoolLeaver = as.integer(lifetime_above_school_leaver)
) %>%
    select(NarrowFieldOfStudyCode, NZQALevel, EarnedAboveSchoolLeaver) %>%
    left_join(study_requirements) %>%
    mutate(StudyRequirementsKey = ifelse(is.na(StudyRequirementsKey), -1, StudyRequirementsKey)) 
head(narrow_field_of_study_level, 20)
tail(narrow_field_of_study_level)

```

narrow_field_of_study_level <- annual_income_table %>% select(NarrowFieldOfStudyCode, NZQALevel) %>% distinct() %>%
    left_join(study_requirements) %>%
    left_join(lifetime_earnings_data) %>%
    mutate(StudyRequirementsKey = ifelse(is.na(StudyRequirementsKey), -1, StudyRequirementsKey)) 
head(narrow_field_of_study_level, 20)
tail(narrow_field_of_study_level)



Ensure that all codes written have already been written to narrow field of study table.
```{r}
dbWriteTable(
    go_web_con, 
    name = "NarrowFieldOfStudyLevel", 
    narrow_field_of_study_level[narrow_field_of_study_level$NarrowFieldOfStudyCode %in% field_of_study_table$NarrowFieldOfStudyCode,] %>%
        filter(!is.na(NZQALevel)), 
    append = TRUE)
```

## Populate Annual Income table

```{r}
fulltime_30_39_income <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_3_avg_income_30_39")) %>% mutate(EmploymentTypeKey = 1L, AgeBandKey = 1L)
head(fulltime_30_39_income)
```

```{r}
fulltime_30_39_income_all <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_avg_3_income_30_39_all"))  %>% mutate(EmploymentTypeKey = 2L, AgeBandKey = 1L)
head(fulltime_30_39_income_all)
```

```{r}
income_30_39 <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "cen_annual_3_avg_income_30_39")) %>% 
    mutate(AgeBandKey = 1L) %>% mutate(PeopleCount = as.integer(people_count), Income = as.integer(mean_income)) %>%
    select(-c(mean_income, people_count))
income_sd <- dbReadTable(go_analysis_con, Id(schema = "lng", table = "cen_inc_age_sd_annual")) %>%
    mutate(AgeBandKey = 2L) %>% mutate(PeopleCount = as.integer(people_count), Income = as.integer(age_sd_income)) %>%
    select(-c(age_sd_income, people_count))
head(income_30_39)
head(income_sd)
```

```{r}
annual_income_table <- income_30_39 %>% rbind(income_sd) %>% 
    mutate(
        NZQALevel = strtoi(level_code, base = 10L),
        NarrowFieldOfStudyCode = substring(Field_Text, 1, 4),
        EmploymentTypeKey = as.integer((employed_code))
    ) %>%
    select(
        NarrowFieldOfStudyCode,
        NZQALevel,
        EmploymentTypeKey,
        AgeBandKey,
        Income,
        PeopleCount
    )

head(annual_income_table)
```

```{r}
dbWriteTable(go_web_con, name = "AnnualIncome", annual_income_table, append = TRUE)
```

# Populate ANZSCO reference tables
```{r}
anzscoL1 <- read_excel("../metadata/AZNSCO_Level1.xlsx", skip = 7) %>% rename(Level1ANZSCOCode = Code, ClassificationTitle = Descriptor)
head(anzscoL1)
dbWriteTable(go_web_con, name = "ANZSCOOccupationL1", anzscoL1, append = TRUE)
```

```{r}
anzscoL2 <- read_excel("../metadata/AZNSCO_Level2.xlsx", skip = 7) %>% rename(Level2ANZSCOCode = Code, ClassificationTitle = Descriptor) %>% mutate(Level1ANZSCOCode = substring(Level2ANZSCOCode, 1,1))
head(anzscoL2)
dbWriteTable(go_web_con, name = "ANZSCOOccupationL2", anzscoL2, append = TRUE)
```

```{r}
anzscoL3 <- read_excel("../metadata/AZNSCO_Level3.xlsx", skip = 7) %>% rename(Level3ANZSCOCode = Code, ClassificationTitle = Descriptor) %>% mutate(Level2ANZSCOCode = substring(Level3ANZSCOCode, 1,2))
head(anzscoL3)
dbWriteTable(go_web_con, name = "ANZSCOOccupationL3", anzscoL3, append = TRUE)
```

```{r}
anzscoL4 <- read_excel("../metadata/AZNSCO_Level4.xlsx", skip = 7) %>% rename(Level4ANZSCOCode = Code, ClassificationTitle = Descriptor) %>% mutate(Level3ANZSCOCode = substring(Level4ANZSCOCode, 1,3))
head(anzscoL4)
dbWriteTable(go_web_con, name = "ANZSCOOccupationL4", anzscoL4, append = TRUE)
```

## Populate ANZSCO L5 table

Need: people count & typical study level.

`public enum TypicalStudyLevel
{
    School = 1, CertDipl = 2, Mixed = 3, Tertiary = 4, Degree = 5
}`

```{r}
#not_working <- tibble(Level5ANZSCOCode = c('999003', '999004'), ClassificationTitle = c('Unemployed', 'Not in Labour Force'))
anzscoL5 <- read_excel("../metadata/ANZSCO_Level5.xlsx", skip = 7) %>% 
    rename(Level5ANZSCOCode = Code, ClassificationTitle = Descriptor) %>% 
 #   rbind(not_working) %>%
    mutate(
        Level1ANZSCOCode = substring(Level5ANZSCOCode, 1,1),
        Level2ANZSCOCode = substring(Level5ANZSCOCode, 1,2),
        Level3ANZSCOCode = substring(Level5ANZSCOCode, 1,3),
        Level4ANZSCOCode = substring(Level5ANZSCOCode, 1,4)
        
    )
head(anzscoL5)
```

```{r}
typical_study_levels <- dbReadTable(go_analysis_con, name = "typical_study_levels") %>%
    rename(TypicalStudyLevelKey = typical_study, Level5ANZSCOCode = i_occupation_code) %>%
    mutate(Level5ANZSCOCode = as.character(Level5ANZSCOCode))
head(typical_study_levels)
```

```{r}
retired_codes <- anzscoL5 %>% filter(ClassificationTitle == "Code retired")
head(retired_codes)
```

```{r}
occupation_table_a <- read_excel("../input_data/t3 - Occupation by field of study - Checked.xlsx", sheet="Occupation totals", skip = 4, col_types = c('text', 'numeric')) %>%
    rename(Level5ANZSCOCode = i_occupation_code, TotalPeopleCount = people_count) %>% right_join(anzscoL5) %>% left_join(typical_study_levels) %>%
    filter(ClassificationTitle != "Code retired")
head(occupation_table_a)
```

```{r}
# leave typical study level for later
dbWriteTable(go_web_con, name = "ANZSCOOccupationL5", occupation_table_a, append = TRUE)
```

# Populate vocational pathways

## Populate reference table
```{r}
vocational_pathway_table <- data.frame(
    VocationalPathwayId = 1:6,
    PathwayName = c("Construction and Infrastructure", "Manufacturing and Technology", "Primary Industries", 
                           "Service Industries", "Social and Community Services", "Creative Industries"),
    PathwayLink = c(
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-construction-and-infrastructure-industries/", 
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-manufacturing-and-technology-industries/",
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-the-primary-industries/",
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-the-service-industries/",
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-social-and-community-service-industries/",
        "https://youthguarantee.education.govt.nz/initiatives/vocational-pathways/learning-and-working-in-the-creative-industries/"
    ),
    stringsAsFactors = FALSE

)
vocational_pathway_table
```

```{r}
dbWriteTable(go_web_con, name = "VocationalPathway", vocational_pathway_table, append = TRUE)
```

## Load vocational pathway to occupation mappings

```{r}
vocational_pathway_map_wide = read_excel("../metadata/VocationalPathwayMap.xlsx") %>% select(-c("ANZSCO group code", "X__1"))
head(vocational_pathway_map_wide)
```

```{r}
vocational_pathway_map_long = vocational_pathway_map_wide %>% pivot_longer(cols = -c(Title, ANZSCO_code), names_to = "pathway_text", values_to = "is_pathway") %>%
    mutate(is_pathway = !is.na(is_pathway)) %>% filter(is_pathway, !ANZSCO_code %in% retired_codes$Level5ANZSCOCode)

head(vocational_pathway_map_long)
```

```{r}
v_pathway2 = vocational_pathway_map_long %>% left_join (vocational_pathway_table, by=c("pathway_text" = "PathwayName"))
head(v_pathway2)
```

```{r}
v_pathway_table = v_pathway2 %>% rename(Level5ANZSCOCode = ANZSCO_code) %>% select(Level5ANZSCOCode, VocationalPathwayId)
head(v_pathway_table)
```

```{r}
dbWriteTable(go_web_con, name = "ANZSCOOccupationL5VocationalPathway", v_pathway_table, append = TRUE)
```

## Populate Careers NZ table


```{r}
career_table = read_csv("../datasets/anzsco_careers_links.csv") %>% rename(Level5ANZSCOCode = Code, CareerLink = Link) %>%
    mutate(CareerLinkEnabled = TRUE) %>% select(Level5ANZSCOCode, Description, CareerLinkEnabled, CareerLink) %>% filter(!is.na(CareerLink))
head(career_table)
```

```{r}
occuapations = dbReadTable(go_web_con, name="ANZSCOOccupationL5")
head(occuapations)
```

```{r}
# temp untill new data avrives
career_table2 = career_table %>% mutate(Level5ANZSCOCode = as.integer(Level5ANZSCOCode)) %>% filter(Level5ANZSCOCode %in% occuapations$Level5ANZSCOCode)
```

```{r}
dbWriteTable(go_web_con, name = "CareersNZInfo", career_table2, append = TRUE)
```


# Populate ANZSCOOccupationL5NarrowFieldOfStudy

```{r}
grad_outcomes_datacube <- read_excel("../input_data/t3b - Occupation by narrow field of study- checked.xlsx", sheet = "Occupation by narrow study", skip=4)
head(grad_outcomes_datacube)
```



```{r}
grad_outcomes_datacube2 <- grad_outcomes_datacube %>% rename(
    Level5ANZSCOCode = i_occupation_code,
    NarrowFieldOfStudyCode = narrow_field_code,
    NZQALevel = i_standard_highest_qual_code,
    People = people_count
)
head(grad_outcomes_datacube2)
```

```{r}
# remove residuals
grad_outcomes_datacube3 <- grad_outcomes_datacube2 %>% 
    filter(
        !NZQALevel %in% c(97, 99),
        NarrowFieldOfStudyCode < 3333
    )
head(grad_outcomes_datacube3)
```

```{r}
dbWriteTable(
    go_web_con, 
    name = "ANZSCOOccupationL5NarrowFieldOfStudy", 
    grad_outcomes_datacube3,
    append = TRUE
)
```

## Populate alternative occupation field of study table
REQUIRES more data from datalab

```{r}
warning("Alternative occupation field of study table not loaded as data is not yet extracted!")
```



