---
title: "Screen Scrape Fees, Durations and SAC funding"
output: html_notebook
---
This code screen-scrapes the TEC and Careers websites

```{r}
library(reticulate)
```


# Commands to setup python in R notebooks

When you run the first python chunk, R will ask to install Miniconda. I would recommend that you do this.
https://docs.conda.io/en/latest/miniconda.html 
Once Minicoda is installed, open (from the start menu) 'Anaconda Powershell Prompt (R-Mini)' and run the following command
`conda activate r-reticulate`
followed by the install commands.

Run these commands on the command line to install

python -m pip install -U selenium
python -m pip install -U bs4
python -m pip install -U pandas

Required for TEC SAC FUNDING webscrap:
python -m pip install -U lxml 
python -m pip install -U xlrd


# Check what version of python is installed, this should work on any version of 3.*
```{python}
import sys
sys.version
```


# Setup
## Load libararies
```{python}
from selenium import webdriver
from bs4 import BeautifulSoup
import pandas as pd
import pickle
import re
```

# GET SAC FUNDING RATES:

This process is more robust than the Careers NZ screen-scraping becuase all the required information is on a single page, in one table.

## Start up webdriver and navigate to starting page

If it breaks at this point you might need to obtain a copy of Chrome webdriver, The error message may give you instructions to install it, you could follow them.

You can go to: https://chromedriver.chromium.org/downloads to download the webdriver. You will also need to have a somewhat up-to-date version of Chrome installed. 
You will need to either:

* include the ChromeDriver location in your PATH environment variable, or
* include the path to ChromeDriver when instantiating webdriver.Chrome (`webdriver.Chrome('/path/to/chromedriver')`)


```{python}
driver = webdriver.Chrome()
driver.get("https://www.tec.govt.nz/funding/funding-and-performance/funding/fund-finder/student-achievement-component-provision-at-level-3-and-above-on-the-nzqf-fund/sac-funding-rates/")
```


```{python}
sac_accordions = driver.find_elements_by_css_selector("div .accordion-item")
sac_accordions
filtered_ob = filter(lambda x: x.find_element_by_class_name("accordion-header").text[0:4] == '2018', sac_accordions)
filtered_ob
filtered_list = list(filtered_ob)
filtered_list
sac_html = filtered_list[0].find_element_by_tag_name("table").get_attribute('outerHTML')
sac_html
```

```{python}
sac_pd = pd.read_html(sac_html)[0]
sac_pd.head()
```


```{python}
def sift_values(row):
    if(len(row[0]) > 1):
        print("move this")
        return row.shift(1)
    else:
        return row

sac_funding2 = sac_pd[3:].apply(sift_values, axis = 1)
sac_funding2
```

```{python}
sac_funding_a = sac_funding2[sac_funding2[1].str.contains("Not valid for PTEs")].shift(-1,axis = 1)
sac_funding_a[0] = sac_funding_a[0].str.slice(0,1)
sac_funding_a
```

```{python}
sac_funding3 = sac_funding2[pd.notna(sac_funding2[0])].append(sac_funding_a)
sac_funding3.columns = ["category", "classification", "1", "2", "3", "4", "5"]
sac_funding3
```

```{python}
sac_funding_rates = sac_funding3.set_index("category").drop(columns=["classification"])
sac_funding_rates
```

```{python}
def convert_to_num(v):
    if(pd.notna(v)):
        return pd.to_numeric(v[1:].replace(",", ""))
    else:
        return v

sac_funding_rates2 = sac_funding_rates.apply(lambda x: x.apply(convert_to_num))
sac_funding_rates2
```

# Map to NZSCED

```{python}
nzsced_map1 = pd.read_excel("../program_costs/NZSCED to SAC.xls")
nzsced_map1.head()
```

```{python}
nzsced_map1[nzsced_map1['NZSCED to SAC: SAC category and level'] == '0601 Medical Studies']
# cells that were na are now NaN
```
```{python}
def get_sac_funding_rate(code, number):
    temp = sac_funding_rates2[number][code]
    if temp is None:
        return NaN
    elif isinstance(temp, float):
       return temp
    else:
        return sac_funding_rates2[number][code][0]
```


```{python}
sac_funding_rates2["3"]["H"]
get_sac_funding_rate("H", "3")
```


```{python}
def use_funding_code(x):
    if pd.isna(x):
        return x
    if len(x) == 2:
        return get_sac_funding_rate(x[0], x[1])
    elif len(x) == 4:
        x = x.split("/")
        cat = x[0][0]
        l1 = x[0][1]
        l2 = x[1]
        rate1 = get_sac_funding_rate(cat, l1)
        rate2 = get_sac_funding_rate(cat, l2)
        #print(x)
        return (rate1 + rate2) / 2
        
    elif len(x) == 5:
        x = x.split("/")
        rate1 = get_sac_funding_rate(x[0][0], x[0][1])
        rate2 = get_sac_funding_rate(x[1][0], x[1][1])
        return (rate1 + rate2) / 2
    else:
        return x

use_funding_code("A1"), use_funding_code("A3/4"), use_funding_code("A1/B2"), use_funding_code(float('NaN'))
```

```{python}
def map_nzsced_to_sac(row):
    return row.apply(use_funding_code)
nzsced_sac_rates = nzsced_map1.set_index("NZSCED to SAC: SAC category and level").apply(map_nzsced_to_sac, axis = 1)
nzsced_sac_rates.head()
```

```{python}
nzsced_sac_rates.to_csv("../program_costs/NZSCED_SAC_rates.csv")
```

# ********************************************************
# CAREERS NZ SCREEN SCRAPING of fees and durations
# ********************************************************

Screen must be wide enough for this to work, full desktop view.

Starting new instance of Chrome driver incase it got confused after the above section.
```{python}
driver = webdriver.Chrome()
driver.get("https://www.careers.govt.nz/searchresults?tab=courses")
```


## define functions to do the screen scraping
```{python}
# returns true if there was a page to navigate to, false otherwise.
def navigate_to_next_page(driver):    
    next_button = driver.find_element_by_class_name("next")
    if next_button.get_attribute("href") == "javascript:;":
        # at the end!
        print("Done")
        return False
    else:
        # click next!
        next_button.click()
        return True
        
def getValue(driver, selector):
    try:
        return driver.find_element_by_css_selector(selector).text
    except:
        return None
def getFeeValue(driver, selector):
    value = getValue(driver, selector)
    if value is not None:
        return value.replace("\nTooltip","")
    else:
        return value
        
def get_course_links_from_page(driver):
    def get_link(element):
        return element.get_attribute("href")
    page_links = driver.find_elements_by_css_selector('.course-provider a')
    course_pages = list(map(get_link,page_links))
    return course_pages
get_course_links_from_page(driver)


# returns true there was a page to navigate to, false otherwise.
def navigate_to_next_page(driver):    
    next_button = driver.find_element_by_class_name("next")
    if next_button.get_attribute("href") == "javascript:;":
        # at the end!
        print("Done")
        return False
    else:
        # click next!
        next_button.click()
        return True

```

# Running the web scraping
Webscraping can take time, during the process a screen of chrome will run itself and navigate through pages on the Careers NZ website.
It does not need to be kept in focus, nor does it need to use your mouse pointer (some other automation tools do). Just don't click on any links while it is running and it should hopefully work.

## Web scraping all links to courses on the Careers NZ site
## WARNING: this will take a while
```{python}
driver.get("https://www.careers.govt.nz/searchresults?tab=courses")
course_pages = get_course_links_from_page(driver)

while (navigate_to_next_page(driver)):
    course_pages.extend(get_course_links_from_page(driver))
    
f= open("../datasets/course_pages.pickle","wb")

pickle.dump(course_pages, f)
f.close()
```

```{python}
f= open("../datasets/course_pages.pickle","rb")

course_pages = pickle.load(f)
f.close()
```

## Web scraping details from all the courses pages on the Careers NZ website
## WARNING: this will take even longer (> 2 hours)
```{python}
def getCourseDetails(driver, url):
    course_details = {
        "url": url,
        "qualName": getValue(driver, "h1.details-title"),
        "providerName": getValue(driver, "h2.details-subtitle"),
        "qualLevel": getValue(driver, ".qualification a"),
        "fieldOfStudy": getValue(driver, ".subject-area p"),
        "duration": getValue(driver, ".length-of-course .decimal-spacing"),
        "annualFee": getFeeValue(driver, '#student-fee .student-fees td[data-label="Annual"]'),
        "govtSubsidy": getFeeValue(driver, 'th.government-subsidy + td')
    }
    
    return course_details
course_dataset = pd.DataFrame(columns=["url", "qualName", "providerName", "qualLevel", "fieldOfStudy", "duration", "annualFee", "govtSubsidy"])
for link in course_pages:
    driver.get(link + "?hello=from_unz&screen_scape")
    course_dataset = course_dataset.append(getCourseDetails(driver, link), ignore_index=True)
    
f= open("../datasets/course_dataset.pickle","wb")
pickle.dump(course_dataset, f)
f.close()
```

```{python}
f= open("../datasets/course_dataset.pickle","rb")

course_dataset_a = pickle.load(f)
f.close()
course_dataset_a.head()
```

# performing post-scrape actions
## convert field of study from type object to string and remove any urls that aren't a qualification.
```{python}
course_dataset2 = course_dataset_a
course_dataset2["fieldOfStudy"] = course_dataset2["fieldOfStudy"] .astype('str') 

course_dataset2 = course_dataset2[["qualifications/view" in a for a in course_dataset2['url']]]
print(course_dataset2.head())

```

## do some cleaning of the course dataset
```{python}
def extract_codes_from_url(url):
    codes = re.search("(?:view\/)(\w+)(?:\/)(\d\d\d\d)$",url)
    return {"qualCode": codes.group(1), "provCode": codes.group(2)}

def get_unit(txt):
    if txt == None:
        return None
    return re.search("(?:\d+ )(\w+)", txt).group(1)

def get_value(txt):
    if txt == None:
        return None
    return re.search("(\S+)(?: \w+)", txt).group(1)

def calculate_durations(txt):
    if txt is None:
        return None
    
    v = float(get_value(txt))
    u = get_unit(txt)
    if u == "Weeks":
        d1 = v / 40
        return round(d1 * 2) / 2 # rounded to the nearest half year
    else:
        return v
        
def read_fee(txt):
    if txt is None:
        return None
    elif txt == "N/A":
        return txt
    else:
        return int(re.sub("[\$,]", "",str(txt)))
def apply_conversions(row):
    row["duration_yr"] = calculate_durations(row["duration"])
    codes = extract_codes_from_url(row["url"])
    row["qualCode"] = codes["qualCode"]
    row["provCode"] = codes["provCode"]
    if row["duration_yr"] == 0.5:
        row["fee"] = read_fee(row["annualFee"]) * 2
    else:
        row["fee"] = read_fee(row["annualFee"])
    return row

course_dataset3 = course_dataset2.apply(apply_conversions, 1)

```


```{python}
course_dataset3.to_csv('course_dataset3.csv')
```


## bring the detailed NZSCED classification codes
Old code, not used.

detailed_nzsced = pd.read_excel("../metadata/NZSCED_detailed.xlsx", skiprows=8, dtype = {"Code": str, "Descriptor": str}).rename(columns={"Descriptor":"fieldOfStudy"})
detailed_nzsced["fieldOfStudy"] = detailed_nzsced.fieldOfStudy.apply(lambda x: re.sub("(\s\s+)", " ", x))
detailed_nzsced.head()

narrow_nzsced = pd.read_excel("../metadata/NZSCED_narrow.xlsx", skiprows=8, dtype = {"Code": str, "Descriptor": str}).rename(columns={"Descriptor":"fieldOfStudy"})
narrow_nzsced["fieldOfStudy"] = narrow_nzsced.fieldOfStudy.apply(lambda x: re.sub("(\s\s+)", " ", x))
narrow_nzsced.head()


## Remove entries which have no useful data
```{python}
data4 = course_dataset3[course_dataset3.fee.notnull() & (course_dataset3.fee != "N/A") & (course_dataset3.fee != "N/AN/A")]
data4.columns
data4.head()
```


## Get the NZSCED classification codes

This attempts to match the titles with all classification levels. Starting with detailed and working up to broad. 
```{python}
def read_nzsced_set(excel_file_path):
    nzsced = pd.read_excel(excel_file_path, skiprows=8, dtype = {"Code": str, "Descriptor": str}).rename(columns={"Descriptor":"fieldOfStudy"})
    nzsced["fieldOfStudy"] = nzsced.fieldOfStudy.apply(lambda x: re.sub("(\s\s+)", " ", x))
    print(nzsced.head())
    return nzsced

#note that pd.read_excel no longer supports xlsx files
def add_nzsced_codes(import_nzsced):
    broad_nzsced = read_nzsced_set("../metadata/NZSCED_broad.xls")
    narrow_nzsced = read_nzsced_set("../metadata/NZSCED_narrow.xls")
    detailed_nzsced = read_nzsced_set("../metadata/NZSCED_detailed.xls")
    
    import_nzsced1 = pd.merge(import_nzsced, detailed_nzsced, on="fieldOfStudy", how="left").rename(columns={"Code": "DetailedNZSCED"})
    import_nzsced2 = pd.merge(import_nzsced1, narrow_nzsced, on="fieldOfStudy", how="left").rename(columns={"Code": "NarrowNZSCED"})
    import_nzsced3 = pd.merge(import_nzsced2, broad_nzsced, on="fieldOfStudy", how="left").rename(columns={"Code": "BroadNZSCED"})
    
    def choose_nzsced_level(row):
        if pd.notnull(row["DetailedNZSCED"]):
            row["NZSCED"] = row["DetailedNZSCED"]
            row["NZSCED_Level"] = 3
            row["NZSCED_Narrow"] = row["DetailedNZSCED"][0:4]
        elif pd.notnull(row["NarrowNZSCED"]):
            row["NZSCED"] = row["NarrowNZSCED"]
            row["NZSCED_Level"] = 2
            row["NZSCED_Narrow"] = row["NarrowNZSCED"]
        elif pd.notnull(row["BroadNZSCED"]):
            row["NZSCED"] = row["BroadNZSCED"]
            row["NZSCED_Level"] = 1
        return row
    
    import_nzsced4 = import_nzsced3.apply(choose_nzsced_level, 1).drop(columns=["DetailedNZSCED", "NarrowNZSCED", "BroadNZSCED"])
    
    return import_nzsced4

#import_nzsced = data4.drop(columns="Code")
course_nzsced = add_nzsced_codes(data4)
course_nzsced.head()
```


## Convert the study level text to numbers
NOTE LEVEL 5 AND 6 DIPLOMAS CATEGORISED AS LEVEL 5
```{python}
levels_of_interest = ["Certificate Level 3", "Diploma Level 5", "Diploma Level 6", "Bachelor Degree Level 7", "Bachelor Degree with Honours Level 8", "Masters Degree Level 9", "Doctorate Level 10"]
levels_of_interest

def convert_qual_level(level):
    if level == 'Certificate Level 3':
        return 3
    elif level == 'Diploma Level 5' or level == 'Diploma Level 6':
        return 5
    elif level == 'Bachelor Degree Level 7':
        return 7
    elif level == 'Bachelor Degree with Honours Level 8':
        return 8
    elif level == 'Masters Degree Level 9':
        return 9
    elif level == 'Doctorate Level 10':
        return 10
    else:
        raise ValueError("This level was not expected {}".format(level))
course_nzsced2 = course_nzsced[course_nzsced.qualLevel.isin(levels_of_interest)].copy()
course_nzsced2["qualLevel"] = course_nzsced2.qualLevel.apply(convert_qual_level)
course_nzsced2.head()
set(course_nzsced.qualLevel)
```

## Group by study level and field of study
We do this because we want fees for each field and level of study, whereas the Careers website only has fees by institution and broad field.
The means of fee and durations are the result
```{python}
level_nzsced_fees = course_nzsced2.groupby(["qualLevel", "NZSCED_Narrow"]).mean()
level_nzsced_fees.head()
level_nzsced_fees.shape

```


```{python}
level_nzsced_fees.to_csv("../datasets/level_nzsced_fees.csv")
```

# ***********************************************
# R section: finalise and do some filling of the gaps in the fees and durations, then merge with SAC funding rates
# ***********************************************

```{r}
library(tidyverse)
```

```{r}
fees_and_durations <- read.csv("../datasets/level_nzsced_fees.csv")
head(fees_and_durations)

fees_and_durations["qualLevel"] <- formatC(fees_and_durations$qualLevel, width = 2, flag = "0")
fees_and_durations["NZSCED_Narrow"] <- formatC(fees_and_durations$NZSCED_Narrow, width = 4, flag = "0")
fees_and_durations["duration_whole_yr"] <- round(fees_and_durations$duration_yr)
head(fees_and_durations)
```


## Use sensible values to full gaps
```{r}
fees_full <- tibble(
    level_code = c('04', '05', '07', '08', '09', '10'),
    fee = c(5500, 6000, 6500, 7000, 8000, 9000),
    duration_yr = c(0.5, 1, 3, 1, 2, 3)
)
fees_full
```


### Get all the level-field combinations

This uses a dataset from the analysis process, do I want to do this?
```{r}
field_narrow <- readxl::read_excel("../metadata/NZSCED_narrow_cen.xlsx", na = "NA", skip=7) %>% filter(Code != '0000')
desired_combinations <- merge(field_narrow["Code"], fees_full["level_code"], by=NULL) %>% rename(field_code = Code)
head(desired_combinations)
```


### Use the fuller
```{r}
fees_and_durations2 <- fees_and_durations %>% rename(level_code = qualLevel, field_code = NZSCED_Narrow) %>%
    right_join(desired_combinations) %>% full_join(by='level_code', fees_full, suffix = c('_c', '_e')) %>%
    mutate(
        duration_yr = ifelse(is.na(duration_yr_c), duration_yr_e, duration_yr_c),
        fee = ifelse(is.na(fee_c), fee_e, fee_c)
    ) %>%
    select(level_code, field_code, duration_yr, fee)
head(fees_and_durations2)

```



## calculate fees for course and likely loan balance
Cumulative duration created which considers previous study for postgraduate qualifications 
NB: PhD cumulative duration = Bach + Mast + PhD
Using median non-fee component for 2018 of $6983 (Table 4, Student Loan Scheme Annual Report 2019)

```{r}
fees_and_durations2$cum_duration = fees_and_durations2$duration_yr
fees_and_durations2 = fees_and_durations2 %>% arrange(field_code, level_code)

subset1 = filter(fees_and_durations2, level_code %in% c('07', '08')) %>% group_by(field_code) %>% mutate(cum_duration = cumsum(duration_yr)) %>% filter(level_code != '07')
subset2 = filter(fees_and_durations2, !level_code %in% c('04', '05', '08')) %>% group_by(field_code) %>% mutate(cum_duration = cumsum(duration_yr))
fees2 = rbind(subset1, subset2, filter(fees_and_durations2,  level_code %in% c('04', '05')))

#Accounting for Honours quals that include Bachelors
fees2[which(fees2$level_code == '08' & fees2$duration_yr >= 4),]$cum_duration = fees2[which(fees2$level_code == '08' & fees2$duration_yr >= 4),]$duration_yr

fees2 = mutate(fees2, 
        fee_component = fee * duration_yr,
        none_fee_component = 6983 * duration_yr 
        #$4800 ~ median living costs + CRC borrowed in 2013 (6983 in 2018)
    ) %>% mutate(
        duration_whole_yr = signif(cum_duration, 1),
        loan_for_course = fee_component + none_fee_component
    )

#Checking
filter(fees2, field_code == '1007') %>% arrange(level_code) %>% View()
```


loan_at_graduation considers loan of previous study for postgrad qualifications
Need to carefully note whether Honours degrees are 1 year or 4+ years.

New method below

```{r}
fees2$loan_at_graduation = fees2$loan_for_course

fees2 = fees2 %>% arrange(field_code, level_code)

subset1 = filter(fees2, level_code %in% c('07', '08')) %>% group_by(field_code) %>% mutate(loan_at_graduation = cumsum(loan_for_course)) %>% filter(level_code != '07')

subset2 = filter(fees2, !level_code %in% c('04', '05', '08')) %>% group_by(field_code) %>% mutate(loan_at_graduation = cumsum(loan_for_course))

costs = rbind(subset1, subset2, filter(fees2,  level_code %in% c('04', '05')))

#Accounting for Honours quals that include Bachelors
costs[which(costs$level_code == '08' & costs$duration_yr >= 4),]$loan_at_graduation = costs[which(costs$level_code == '08' & costs$duration_yr >= 4),]$loan_for_course

#Checking
filter(costs, field_code == '0401') %>% arrange(level_code) %>% View()
```

```{r}
# costs <- fees2 %>% mutate(
#     loan_at_graduation = ifelse(level_code %in% c("03", "04", "05", "07"), 
#                           loan_for_course,
#                               ifelse(level_code %in% c("08", "09"), 
#                                   loan_for_course + fees2[fees2$level_code == "07" & fees2$field_code == field_code,]$loan_for_course,
#                                   #Else, Level 10 PhD
#                                   loan_for_course + 
#                                      fees2[fees2$level_code == "07" & fees2$field_code == field_code,"loan_for_course"] +
#                                      fees2[fees2$level_code == "09" & fees2$field_code == field_code,"loan_for_course"]
#                               )
#                          ),
# )
# 

```

## Merge in SAC rates
```{r}
sac_funding <- read.csv("../program_costs/NZSCED_SAC_rates.csv")
print(colnames(sac_funding))

colnames(sac_funding) <- c('field_text', 'lvl 04', 'lvl 05', 'lvl 07', 'lvl 08', 'lvl 09', 'lvl 10')
print(colnames(sac_funding))
```

```{r}
sac_funding2 <- sac_funding %>% mutate(field_code = substr(field_text, 1, 4)) %>% 
    select(-c(field_text)) %>% 
    pivot_longer(cols = -c(field_code), names_to = "level_code", names_prefix = "lvl ", values_to = "sac_funding")
tail(sac_funding2)
```

```{r}
sac_paid <- costs %>% left_join(sac_funding2, by=c("level_code", "field_code")) %>%
    mutate(
        total_funding = sac_funding * duration_yr
    )
sac_paid %>% filter(field_code == '0607') %>% head(8)
```

## Save fee, duration and funding dataset
```{r}
write.csv(sac_paid, "../program_costs/fees_durations_SAC_NEW.csv", row.names = FALSE)
```

